\documentclass{beamer}
\usepackage[utf8]{inputenc}
\usepackage{graphicx}
\author[Sowmya Vajjala]{Instructor: Sowmya Vajjala}

\title[LING 410]{LING 410: Language as Data}
\subtitle{Semester: Spring '18}

\date{26 Apr 2018}

\institute{Iowa State University, USA}
%%%%%%%%%%%%%%%%%%%%%%%%%%%

\begin{document}

\begin{frame}\titlepage
\end{frame}

\begin{frame}
\frametitle{To Do}
\begin{itemize}
\item Upload your presentations on Canvas by today. I cannot grade you until you submit. One submission per group is fine. 
\item Guidelines for final submission:
\begin{itemize}
\item Deadline: May 3, end of the day (15\% of the final grade)
\item What it should contain: All code you wrote, any visuals you generated, and a report (not exceeding 5 pages) describing your research question, what you did with R, what is the data, what conclusions you can draw from your results etc.
\end{itemize}
\item Submit your course evals - do them today in the remaining time!!
\end{itemize}
\end{frame}

\begin{frame}
\frametitle{Quick Summary of what we learnt}
In terms of working with textual data, we looked at:
\begin{itemize}
\item How to work with different formats/sources of data (txt, html, tweets, guardian/times etc)
\item How to calculate some basic statistics (frequencies, lexical richness etc)
\item How to look for words in context (KWIC, ngrams)
\item How to create some basic plots and visualizations (plot function, stylo, wordclouds etc)
\item How to work with text classification, clustering, and topic modeling
\end{itemize}
\end{frame}

\begin{frame}
\frametitle{Far from complete ..}
Remember: We did only a basics course.
\begin{itemize}
\item There is lot more to explore, and there are more things you can do for your discipline's research problems or in jobs, with this knowledge.
\item I hope you will continue being interested on this and will enhance your knowledge in future.
\item It is not easy, but it is certainly doable, with some patience, consistent effort, and with a mentor at your workplace in future.
\item Check: tidytextmining.com 
\end{itemize}
\end{frame}


\begin{frame}
\frametitle{Good luck!}
\begin{itemize}
\item Have fun with data and good luck for your future!
\item My email: sowmya@iastate.edu, vbsowmya@gmail.com 
\item If you need a reference in future, you can contact me (there is no guarantee I will write a letter for each person who asks me, though!)
\end{itemize}
\end{frame}

\begin{frame}
\frametitle{}
\centering
\Large It is a challenging course, and you are all reaching the finish line. So, Congratulations!!
\end{frame}

\end{document}
