\documentclass[11pt,a4paper]{article}
\usepackage{url}

\begin{document}
\begin{center}
  Spring Semester 2018 \\ Iowa State University\\[3ex]
  {\large LING 410X - Language as Data}\\[3ex]
  \textbf{Assignment 1} \\ \textbf{Submission Deadline: 27 Jan 2018, end of the day}
\end{center}


\paragraph{Instructions:} This assignment consists of two questions. Each question carries 5 marks. Upload your submission as a zip file in your\_Lastname\_A1.zip format. If any of the programs does not run and throws errors, you cannot get credit for that. Late submissions are allowed, but will not be awarded full credit.

\section*{Question 1} 
Pick one of the following disciplines that is the closest to your major or the one that interests you the most:
\begin{itemize}
\item Psychology
\item Sociology
\item Literature
\item Political Science
\item Linguistics
\item Business, Management and Finance
\item Life sciences
\item Agronomy and Agriculture
\item Information technology
\end{itemize}
Now, do a little bit of research and list 5 problems where text processing is useful in these areas. Give a list of stuff you read (news articles, wikipedia etc) as references at the end of your report. Your report should be a pdf file, can be upto 3 pages long, and without any double spacing.

\section*{Question 2}
Read the tutorial pdf I am providing with this document (A1-HelpFile.pdf), and answer the following questions. 

Consider this paragraph from ISU website on Cyclone Athletics: \\
\textit{For only the second time in its history, the Iowa State Athletics Department finished in the Top 40 of the nation's most-successful athletics departments from a competitive standpoint. The 2014 Learfield Sports Directors' Cup final tally has the Cyclones in 38th place with 585.75 points. The school's all-time best finish is 34th in 2010. Iowa State had the fifth-best finish among Big 12 schools and tops in the state of Iowa. The school earned 160 points in the spring season, which equaled its all-time record. The women’s track \& field (60), women’s golf (51) and men’s golf (49) accounted for the spring points. ISU had averaged only 46 points in the spring season previously. Iowa State also earned points from wrestling (64.5), men's basketball (64), women's cross country (63), gymnastics (59.25), women's indoor track \& field (51.5), men's indoor track \& field (46.5), men's cross country (27), women's basketball (25) and volleyball (25).}

Based on what you learnt in the tutorial, answer the following questions using R. Submit the lines you typed in R along with the output you got as a pdf file. 
\begin{enumerate}
\item Read this paragraph into a string variable called cyclones.
\item What are the number of characters in cyclones?
%nchar(cyclones)
\item Uppercase cyclones and store it in another variable cyclones\_upper
%cyclones_lower <- tolower(cyclones)
\item Substitute all occurrences of numbers of the form XY.Z (i.e., numbers with only 1 digit after decimal) with NUM and show the output.
 %gsub("[0-9][0-9]", "NUM", cyclones_lower)
\item Split cyclones text wherever a period (.) occurs and print the output. What is the type of output you get (string, list, numbers etc)? 
%strsplit(cyclones,"\\.")
\item Convert cyclones text into title case and print the output. What does the output look like? 
\end{enumerate}

\end{document}

